---
title: "Schema"
output: html_document
date: "2024-03-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(DBI)
library(RSQLite)

#install.packages("DBI")
#install.packages("RSQLite")
```


##Connect to database
```{r}
connect <- dbConnect(RSQLite::SQLite(), "database.db")
```


## Create CUSTOMER Table #Already run
```{sql connection=connect}
CREATE TABLE 'CUSTOMERS' (
  'customer_id' INT PRIMARY KEY, 
  'first_name' VARCHAR(250) NOT NULL,
  'last_name' VARCHAR(250) NOT NULL,
  'email' VARCHAR(250) NOT NULL,
  'mobile_no' INT NOT NULL,
  'street' VARCHAR (50) NOT NULL,
  'city' TEXT NOT NULL,
  'state' TEXT NOT NULL,
  'country' TEXT NOT NULL,
  'postcode' VARCHAR (50) NOT NULL
) ;  

```

## Create PRODUCT table
```{sql connection=connect}
CREATE TABLE 'PRODUCT' (
  'product_id' INT PRIMARY KEY, 
  'product_name' VARCHAR(250) NOT NULL,
  'product_description' VARCHAR(250),
  'unit_price' VARCHAR(250) NOT NULL,
  'stock_on_hand' INT NOT NULL,
  'main_product_id' INT,
  'category_id' INT NOT NULL,
  'supplier_id' INT NOT NULL,
  FOREIGN KEY (supplier_id)
    REFERENCES SUPPLIER (supplier_id),
  FOREIGN KEY (category_id)
    REFERENCES CATEGORY (category_id)
) ;  

```



##Create ORDER table for CUSTOMER-PRODUCT relationship
```{sql connection=connect}
CREATE TABLE 'ORDER' (
  'order_id' INT PRIMARY KEY,
  'customer_id' INT, 
  'product_id' INT,
  'order_date' DATE NOT NULL,
  'order_status' TEXT NOT NULL, 
  'order_quantity' INT NOT NULL,
  'promo_code' VARCHAR(10),
  'discount_percent' NUMERIC DEFAULT 0,
  'payment_detail' TEXT NOT NULL,
  'delivery_fee' NUMERIC DEFAULT 0,
  FOREIGN KEY (customer_id)
    REFERENCES CUSTOMER (customer_id),
  FOREIGN KEY (product_id)
    REFERENCES PRODUCT (product_id)
) ;  

```


## Create ADVERTISEMENTS table
```{sql connection=connect}
CREATE TABLE 'ADVERTISEMENTS' (
  'ad_id' INT PRIMARY KEY, 
  'ad_number' INT NOT NULL,
  'ad_price' NUMERIC NOT NULL, 
  'ad_place' VARCHAR (50) NOT NULL
) ;

```


##Create ADVERTISE_IN table for PRODUCT-ADVERTISEMENT relationship
```{sql connection=connect}
CREATE TABLE 'ADVERTISE_IN' (
  'product_id' INT PRIMARY KEY, 
  'ad_id' INT PRIMARY KEY, 
  FOREIGN KEY (product_id)
    REFERENCES PRODUCT (product_id),
  FOREIGN KEY (ad_id)
    REFERENCES ADVERTISEMENTS (ad_id)
) ;

```

## Create SUPPLIER table
```{sql connection=connect}
CREATE TABLE 'SUPPLIER' (
  'supplier_id' INT PRIMARY KEY, 
  'first_name' VARCHAR(250) NOT NULL,
  'last_name' VARCHAR(250) NOT NULL, 
  'email' VARCHAR(250) NOT NULL,
  'mobile_no' INT NOT NULL,
) ;  

```

## Create CATEGORY table
```{sql connection=connect}
CREATE TABLE 'CATEGORY' (
  'category_id' INT PRIMARY KEY, 
  'category_name' VARCHAR(250) NOT NULL
) ;

```


#Import to each Entity
## Loading using ``read_csv`` and ``dbWriteTable()``


##This is not our version, I just copy it from the workshop 
```{r dataloading,message=FALSE,warning=FALSE}
PurchasesDec <- readr::read_csv("bibitor/PurchasesFINAL12312016a.csv")
BegInvDec <- readr::read_csv("bibitor/BegInvFINAL12312016a.csv")
EndInvDec <- readr::read_csv("bibitor/EndInvFINAL12312016a.csv")

# in a similar way we can read and load the other tables and give the names  
# that are in the PWC data case study - see slide 9 from the uploaded pdf
VendorInvoicesDec <- readr::read_csv("bibitor/InvoicePurchases12312016a.csv")
```

## Write them to the database 
```{r writebacktodb}
RSQLite::dbWriteTable(my_connection,"PurchasesDec",PurchasesDec,overwrite=TRUE)
RSQLite::dbWriteTable(my_connection,"BegInvDec",BegInvDec,overwrite=TRUE)
RSQLite::dbWriteTable(my_connection,"EndInvDec",EndInvDec,overwrite=TRUE)
RSQLite::dbWriteTable(my_connection,"VendorInvoicesDec",VendorInvoicesDec,overwrite=TRUE)

```

