---
title: "Schema"
output: html_document
date: "2024-03-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(DBI)
library(RSQLite)

#install.packages("DBI")
#install.packages("RSQLite")
```

##Adjustment 14/3/2024: delete state (focus only address in the UK), Change datatype of all primary key to varchar, create DISCOUNT table to be in  compliance with 3NF
##Adjustment 15/3/2024: delete SUPPLIER first_name and last_name, and change to supplier_name
## added category fee, product review, rename attributes

##Connect to database

```{r}
connect <- dbConnect(RSQLite::SQLite(), "database.db")
```

## Create CUSTOMER Table 

        
```{sql connection=connect}
CREATE TABLE CUSTOMER (
  customer_id VARCHAR(250) PRIMARY KEY, 
  first_name VARCHAR(250) NOT NULL,
  last_name VARCHAR(250) NOT NULL,
  gender VARCHAR(10),
  customer_email VARCHAR(250) NOT NULL,
  customer_mobile VARCHAR(15) NOT NULL
);
```

## Create ADDRESS table

```{sql connection=connect}
CREATE TABLE ADDRESS (
  postcode VARCHAR(20) PRIMARY KEY, 
  customer_id INT NOT NULL,
  street VARCHAR(50) NOT NULL,
  city VARCHAR(100) NOT NULL,
  country VARCHAR(100) NOT NULL,
  FOREIGN KEY (customer_id)
    REFERENCES CUSTOMER (customer_id)
);
```


## Create PRODUCT table

```{sql connection=connect}
CREATE TABLE PRODUCT (
  product_id VARCHAR(250) PRIMARY KEY, 
  product_name VARCHAR(250) NOT NULL,
  product_description VARCHAR(250),
  product_rating DECIMAL(5,2),
  unit_price DECIMAL(10,2) NOT NULL,
  stock_on_hand INT NOT NULL,
  main_product_id INT,
  category_id INT NOT NULL,
  supplier_id INT NOT NULL,
  FOREIGN KEY (supplier_id)
    REFERENCES SUPPLIER (supplier_id),
  FOREIGN KEY (category_id)
    REFERENCES CATEGORY (category_id)
);
```

##Create DISCOUNT table

````{sql connection=connect}
CREATE TABLE DISCOUNT (
  promo_code VARCHAR(20) PRIMARY KEY, 
  discount_amount INT NOT NULL
);
```



##Create ORDER table for CUSTOMER-PRODUCT relationship
```{sql connection=connect}
CREATE TABLE ORDER_ITEM (
  order_id VARCHAR(250) PRIMARY KEY,
  product_id VARCHAR(250), 
  order_quantity INT NOT NULL,
  FOREIGN KEY (product_id)
    REFERENCES PRODUCT (product_id)
);

```


```{sql connection=connect}
CREATE TABLE ORDER_DETAIL (
  order_id VARCHAR(250) PRIMARY KEY,
  customer_id VARCHAR(250), 
  order_date DATE NOT NULL,
  order_status VARCHAR(50) NOT NULL, 
  promo_code VARCHAR(20),
  payment_method TEXT NOT NULL,
  FOREIGN KEY (customer_id)
    REFERENCES CUSTOMER (customer_id),
  FOREIGN KEY (order_id)
    REFERENCES CUSTOMER (customer_id),
  FOREIGN KEY (promo_code)
    REFERENCES DISCOUNT (promo_code)
);
```


## Create ADVERTISEMENT table

```{sql connection=connect}
CREATE TABLE ADVERTISEMENT (
  ad_id VARCHAR(250) PRIMARY KEY, 
  ad_frequency INT NOT NULL,
  ad_place VARCHAR(50) NOT NULL,
  ad_price DECIMAL(10, 2) NOT NULL, 
);
```

##Create ADVERTISE_IN table for PRODUCT-ADVERTISEMENT relationship

```{sql connection=connect}
CREATE TABLE ADVERTISE_IN (
  product_id VARCHAR(250),
  ad_id VARCHAR(250),
  PRIMARY KEY (product_id, ad_id),
  FOREIGN KEY (product_id) REFERENCES PRODUCT (product_id),
  FOREIGN KEY (ad_id) REFERENCES ADVERTISEMENTS (ad_id)
);
```

## Create SUPPLIER table

```{sql connection=connect}
CREATE TABLE SUPPLIER (
  supplier_id VARCHAR(250) PRIMARY KEY, 
  supplier_name VARCHAR(250) NOT NULL,
  supplier_email VARCHAR(250) NOT NULL,
  supplier_mobile VARCHAR(20) NOT NULL
);
```

## Create CATEGORY table

```{sql connection=connect}
CREATE TABLE CATEGORY (
  category_id VARCHAR(250) PRIMARY KEY, 
  category_name VARCHAR(250) NOT NULL,
  category_fee INT NOT NULL,
);
```

#Import to each Entity \## Loading using `read_csv` and `dbWriteTable()`

##This is not our version, I just copy it from the workshop

```{r dataloading,message=FALSE,warning=FALSE}
PurchasesDec <- readr::read_csv("bibitor/PurchasesFINAL12312016a.csv")
BegInvDec <- readr::read_csv("bibitor/BegInvFINAL12312016a.csv")
EndInvDec <- readr::read_csv("bibitor/EndInvFINAL12312016a.csv")

# in a similar way we can read and load the other tables and give the names  
# that are in the PWC data case study - see slide 9 from the uploaded pdf
VendorInvoicesDec <- readr::read_csv("bibitor/InvoicePurchases12312016a.csv")
```

## Write them to the database

```{r writebacktodb}
RSQLite::dbWriteTable(my_connection,"PurchasesDec",PurchasesDec,overwrite=TRUE)
RSQLite::dbWriteTable(my_connection,"BegInvDec",BegInvDec,overwrite=TRUE)
RSQLite::dbWriteTable(my_connection,"EndInvDec",EndInvDec,overwrite=TRUE)
RSQLite::dbWriteTable(my_connection,"VendorInvoicesDec",VendorInvoicesDec,overwrite=TRUE)

```
